Full-day, in home validation of infant body position measurements from inertial sensors

### 2023-03-15 manuscript.Rmd
figure captions

```{r metrics, warning=FALSE, fig.cap="Metrics", fig.align="center", fig.width=6, fig.height=3, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
metrics <- read_csv("data/group_split_metrics.csv") %>% 
  filter(model %in% c("group", "split")) %>% 
  mutate(Model = factor(model, levels = c("group", "split"), labels = c("Group", "Individual")))

@@

  apa_table(metrics_summary, col.names = c("Metric", "Median", "Mean", "SD", "Median", "Mean", "SD"),    col_spanners = list("Group" = c(2, 4), "Individual" = c(5, 7)),
          caption = "Summary statistics for model performance metrics shown separately for group and individual models.")

@@

apa_table(class_table, col.names = c("Metric","Position", "Median", "Mean", "SD", "Median", "Mean", "SD"), col_spanners = list("Group" = c(3, 5), "Individual" = c(6, 8)),
          midrules = c(5, 10),
          caption = "Model performance metrics for each body position category, shown separately for group and individual models.")
metrics_long_summary <- metrics_long_summary %>% 
  mutate(rowname = paste(Class, metric)) %>% 
  column_to_rownames("rowname")
```

@@

```{r part2overall, warning=FALSE, fig.cap="Overall agreement between human-coded body position and model-predicted body position in the distal comparison. Agreement for group models is shown in (A) and agreement for individual models is shown in (B). Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant. The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=7, fig.height=6.5, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}

@@

#Overall
group_agree_cor <- cor_test(agree$group, vars = c("prop_model", "prop_human")) %>% pull(cor)
split_agree_cor <- cor_test(agree$split, vars = c("prop_model", "prop_human")) %>% pull(cor)
group_agree_cor_noout <- agree$group %>% filter_outliers %>% 
  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
split_agree_cor_noout <- agree$split %>% filter_outliers %>%
  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
overall_row <- tibble("Overall", group_agree_cor, split_agree_cor, group_agree_cor_noout, split_agree_cor_noout)
colnames(overall_row) <- colnames(pt2corr)
pt2corr <- bind_rows(pt2corr, overall_row)
apa_table(pt2corr, col.names = c("Position", "Group", "Individual", "Group", "Individual"), col_spanners = list(`With Outliers` = c(2, 3), `Without Outliers` = c(4, 5)),
          midrules = 5, caption = "Correlations between human-coded and model-predicted body position durations across the entire long delay period. Correlations are provided within each posture and overall, and computed separately using group and individual models with and without outlier participants.")
```

@@

```{r part2bins, warning=FALSE, fig.cap="Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison. Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins. Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute. Performance is plotted separately for (A) group models and (B) individual models. The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=8, fig.height=4.5, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
calculate_agreement_bins <- function(ds) {
  ds <- ds %>% group_by(file) %>% 
    mutate(sample = row_number(), bin = floor(sample/300)) %>% 
    add_count(file, bin) %>% ungroup() %>% filter(n > 210) %>% group_by(file) %>% mutate(num_bins = length(unique(bin))) %>% ungroup()
  model <- ds %>% count(file, bin, num_bins, pos) %>% add_count(file, bin,num_bins,  wt = n, name = "total") %>% mutate(prop_model = n/total)
  human <- ds %>% count(file, bin, num_bins,  code) %>% add_count(file, bin, num_bins, wt = n, name = "total") %>% mutate(prop_human = n/total)
  agree <- full_join(select(model, file, pos, prop_model, bin, num_bins), 
                     select(human, file, code, prop_human, bin, num_bins), 
                     by = c("file" = "file", "pos" = "code", "bin" = "bin", "num_bins" = "num_bins"))
  agree <- agree %>% mutate(prop_human = ifelse(is.na(prop_human), 0, prop_human),
                            prop_model = ifelse(is.na(prop_model), 0, prop_model),
                            pos - factor(pos, levels = pos_levels))
  agree <- complete(agree, nesting(file, bin, num_bins), pos, fill = list(prop_human = 0, prop_model = 0))
}

@@

group_agree_cor <- cor_test(agree_bins$group, vars = c("prop_model", "prop_human")) %>% pull(cor)
split_agree_cor <- cor_test(agree_bins$split, vars = c("prop_model", "prop_human")) %>% pull(cor)
group_agree_cor_noout <- agree_bins$group %>% filter_outliers %>% 
  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
split_agree_cor_noout <- agree_bins$split %>% filter_outliers %>%
  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
overall_row <- tibble("Overall", group_agree_cor, split_agree_cor, group_agree_cor_noout, split_agree_cor_noout)
colnames(overall_row) <- colnames(pt2corr_bins)
pt2corr_bins <- bind_rows(pt2corr_bins, overall_row)
rm(overall_row)
apa_table(pt2corr_bins, col.names = c("Position", "Group", "Individual", "Group", "Individual"), col_spanners = list(`With Outliers` = c(2, 3), `Without Outliers` = c(4, 5)),
          midrules = 5, caption = "Correlations between human-coded and model-predicted body position durations using 10-minute bins during the distal comparison. Correlations are provided within each posture and overall, and computed separately using group and individual models with and without outlier participants.")
```

@@

```{r timelines, warning=FALSE, fig.cap="Timelines", fig.align="center", fig.width=7.75, fig.height=8, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
timeline_data <- read_csv("data/imu-compiled.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")),
         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
  filter(id_uni != 10401) %>% 
  pivot_longer(cols = nap_time:upright_time, names_to = "position", values_to = "prop") %>% 
  mutate(Position = factor(position,
                           levels = c("nap_time","upright_time", "sit_time", "prone_time", "supine_time", "held_time"),
                           labels = c("Nap","Upright", "Sitting", "Prone", "Supine", "Held"))) %>% 
  arrange(Position, age)

@@

```{r age, warning=FALSE, fig.cap="Age trends", fig.align="center", fig.width=7.5, fig.height=7, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
fullday_group <- read_csv("data/imu-compiled.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")),
         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
  filter(id_uni != 10401)
fullday_split <- read_csv("data/imu-compiled-split.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")),
         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
  filter(id_uni != 10401)
fullday <- list("group" = fullday_group, "split" = fullday_split)

>>>>

```{r metrics, warning=FALSE, fig.cap="Metrics of agreement between human-annotated body position and model predictions of body position from the proximal comparison period. Overall accuracy (A) and Cohen's Kappa (B) are plotted separately for group models and individual models. Each blue circle represents that accuracy for each recording session. Horizontal black bars indicates the mean across sessions.", fig.align="center", fig.width=6, fig.height=3, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
metrics <- read_csv("data/group_split_metrics.csv") %>% 
  filter(model %in% c("group", "split")) %>% 
  mutate(Model = factor(model, levels = c("group", "split"), labels = c("Group", "Individual")))

@@

  apa_table(metrics_summary, col.names = c("Metric", "Median", "Mean", "SD", "Median", "Mean", "SD"),    col_spanners = list("Group" = c(2, 4), "Individual" = c(5, 7)),
          caption = "Model performance metrics (overall accuracy, Cohen's Kappa, sensitivity, positive predictive value) from the proximal comparison period. Descriptive statistics are shown separately for group and individual models.")

@@

apa_table(class_table, col.names = c("Metric","Position", "Median", "Mean", "SD", "Median", "Mean", "SD"), col_spanners = list("Group" = c(3, 5), "Individual" = c(6, 8)),
          midrules = c(5, 10),
          caption = "Model performance metrics for each body position (supine, prone, sitting, upright, and held) during the proximal comparison period, shown separately for group and individual models.")
metrics_long_summary <- metrics_long_summary %>% 
  mutate(rowname = paste(Class, metric)) %>% 
  column_to_rownames("rowname")
```

@@

```{r part2overall, warning=FALSE, fig.cap="Overall agreement between human-coded body position and model-predicted body position in the distal comparison. Agreement for group models is shown in (A) and agreement for individual models is shown in (B). Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represents data for a single session. The two outlier participants are plotted in dark gray, with a different shape marking each individual. ", fig.align="center", fig.width=7, fig.height=6.5, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}

@@

#Overall
group_agree_cor <- cor_test(agree$group, vars = c("prop_model", "prop_human")) %>% pull(cor)
split_agree_cor <- cor_test(agree$split, vars = c("prop_model", "prop_human")) %>% pull(cor)
group_agree_cor_noout <- agree$group %>% filter_outliers %>% 
  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
split_agree_cor_noout <- agree$split %>% filter_outliers %>%
  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
overall_row <- tibble("Overall", group_agree_cor, split_agree_cor, group_agree_cor_noout, split_agree_cor_noout)
colnames(overall_row) <- colnames(pt2corr)
pt2corr <- bind_rows(pt2corr, overall_row)
apa_table(pt2corr, col.names = c("Position", "Group", "Individual", "Group", "Individual"), col_spanners = list(`With Outliers` = c(2, 3), `Without Outliers` = c(4, 5)),
          midrules = 5, caption = "Correlations between human-coded and model-predicted body position durations across the entire distal comparison period. Correlations are provided within each body position and overall. Correlations are presented separately for group and individual models with and without the two outlier participants.")
```

@@

```{r part2bins, warning=FALSE, fig.cap="Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison period. Each point shows the mean and SE for a single participant's recording session for each body position, summarizing the prediction difference for each of their 10-minute bins. Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute. Performance is plotted separately for (A) group models and (B) individual models. The two outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=8, fig.height=4.5, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
calculate_agreement_bins <- function(ds) {
  ds <- ds %>% group_by(file) %>% 
    mutate(sample = row_number(), bin = floor(sample/300)) %>% 
    add_count(file, bin) %>% ungroup() %>% filter(n > 210) %>% group_by(file) %>% mutate(num_bins = length(unique(bin))) %>% ungroup()
  model <- ds %>% count(file, bin, num_bins, pos) %>% add_count(file, bin,num_bins,  wt = n, name = "total") %>% mutate(prop_model = n/total)
  human <- ds %>% count(file, bin, num_bins,  code) %>% add_count(file, bin, num_bins, wt = n, name = "total") %>% mutate(prop_human = n/total)
  agree <- full_join(select(model, file, pos, prop_model, bin, num_bins), 
                     select(human, file, code, prop_human, bin, num_bins), 
                     by = c("file" = "file", "pos" = "code", "bin" = "bin", "num_bins" = "num_bins"))
  agree <- agree %>% mutate(prop_human = ifelse(is.na(prop_human), 0, prop_human),
                            prop_model = ifelse(is.na(prop_model), 0, prop_model),
                            pos - factor(pos, levels = pos_levels))
  agree <- complete(agree, nesting(file, bin, num_bins), pos, fill = list(prop_human = 0, prop_model = 0))
}

@@

group_agree_cor <- cor_test(agree_bins$group, vars = c("prop_model", "prop_human")) %>% pull(cor)
split_agree_cor <- cor_test(agree_bins$split, vars = c("prop_model", "prop_human")) %>% pull(cor)
group_agree_cor_noout <- agree_bins$group %>% filter_outliers %>% 
  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
split_agree_cor_noout <- agree_bins$split %>% filter_outliers %>%
  cor_test(vars = c("prop_model", "prop_human")) %>% pull(cor)
overall_row <- tibble("Overall", group_agree_cor, split_agree_cor, group_agree_cor_noout, split_agree_cor_noout)
colnames(overall_row) <- colnames(pt2corr_bins)
pt2corr_bins <- bind_rows(pt2corr_bins, overall_row)
rm(overall_row)
apa_table(pt2corr_bins, col.names = c("Position", "Group", "Individual", "Group", "Individual"), col_spanners = list(`With Outliers` = c(2, 3), `Without Outliers` = c(4, 5)),
          midrules = 5, caption = "Correlations between human-coded and model-predicted body position durations using 10-minute bins during the distal comparison period. Correlations are provided within each posture and overall, and computed separately using group and individual models with and without outlier participants.")
```

@@

```{r timelines, warning=FALSE, fig.cap="Full-day timelines for each individual recording session, split by (A) younger infants and (B) older infants. Each participant's timeline shows a stacked bar graph with the proportion of time spent in each of the body position for every 5-minute period throughout the day, based on group model predictions of body position. The x-axis shows time of day. Caregiver-reported naps are marked by gray bars; blank gaps indicate caregiver-reported times that the sensor garment was removed for diaper changes, baths, or trips out of the house.", fig.align="center", fig.width=7.75, fig.height=8, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
timeline_data <- read_csv("data/imu-compiled.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")),
         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
  filter(id_uni != 10401) %>% 
  pivot_longer(cols = nap_time:upright_time, names_to = "position", values_to = "prop") %>% 
  mutate(Position = factor(position,
                           levels = c("nap_time","upright_time", "sit_time", "prone_time", "supine_time", "held_time"),
                           labels = c("Nap","Upright", "Sitting", "Prone", "Supine", "Held"))) %>% 
  arrange(Position, age)

@@

```{r age, warning=FALSE, fig.cap="Age differences in daily body position predicted from (A) group models and (B) individual models. Each circle represents one full-day recording session's proportion of time in each body position (y-axis) against age in months (x-axis). Black lines indicate best fit regression lines indicating decreases in supine time and increases in sitting and upright time with age.", fig.align="center", fig.width=7.5, fig.height=7, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
fullday_group <- read_csv("data/imu-compiled.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")),
         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
  filter(id_uni != 10401)
fullday_split <- read_csv("data/imu-compiled-split.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")),
         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
  filter(id_uni != 10401)
fullday <- list("group" = fullday_group, "split" = fullday_split)

### 2023-03-15 manuscript.tex
figure captions

\caption{Metrics}\label{fig:metrics}
\end{figure}

@@

\caption{\label{tab:metrics-table}Summary statistics for model performance metrics shown separately for group and individual models.}

@@

\caption{\label{tab:metrics-by-class}Model performance metrics for each body position category, shown separately for group and individual models.}

@@

\caption{Overall agreement between human-coded body position and model-predicted body position in the distal comparison. Agreement for group models is shown in (A) and agreement for individual models is shown in (B). Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant. The three outlier participants are plotted in dark gray, with a different shape marking each individual.}\label{fig:part2overall}
\end{figure}

@@

\caption{\label{tab:pt2overalltable}Correlations between human-coded and model-predicted body position durations across the entire long delay period. Correlations are provided within each posture and overall, and computed separately using group and individual models with and without outlier participants.}

@@

\caption{Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison. Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins. Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute. Performance is plotted separately for (A) group models and (B) individual models. The three outlier participants are plotted in dark gray, with a different shape marking each individual.}\label{fig:part2bins}
\end{figure}

@@

\caption{\label{tab:pt2binstable}Correlations between human-coded and model-predicted body position durations using 10-minute bins during the distal comparison. Correlations are provided within each posture and overall, and computed separately using group and individual models with and without outlier participants.}

@@

\caption{Timelines}\label{fig:timelines}
\end{figure}

@@

\caption{Age trends}\label{fig:age}
\end{figure}

>>>>

\caption{Metrics of agreement between human-annotated body position and model predictions of body position from the proximal comparison period. Overall accuracy (A) and Cohen's Kappa (B) are plotted separately for group models and individual models. Each blue circle represents that accuracy for each recording session. Horizontal black bars indicates the mean across sessions.}\label{fig:metrics}
\end{figure}

@@

\caption{\label{tab:metrics-table}Model performance metrics (overall accuracy, Cohen's Kappa, sensitivity, positive predictive value) from the proximal comparison period. Descriptive statistics are shown separately for group and individual models.}

@@

\caption{\label{tab:metrics-by-class}Model performance metrics for each body position (supine, prone, sitting, upright, and held) during the proximal comparison period, shown separately for group and individual models.}

@@

\caption{Overall agreement between human-coded body position and model-predicted body position in the distal comparison. Agreement for group models is shown in (A) and agreement for individual models is shown in (B). Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represents data for a single session. The two outlier participants are plotted in dark gray, with a different shape marking each individual. }\label{fig:part2overall}
\end{figure}

@@

\caption{\label{tab:pt2overalltable}Correlations between human-coded and model-predicted body position durations across the entire distal comparison period. Correlations are provided within each body position and overall. Correlations are presented separately for group and individual models with and without the two outlier participants.}

@@

\caption{Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison period. Each point shows the mean and SE for a single participant's recording session for each body position, summarizing the prediction difference for each of their 10-minute bins. Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute. Performance is plotted separately for (A) group models and (B) individual models. The two outlier participants are plotted in dark gray, with a different shape marking each individual.}\label{fig:part2bins}
\end{figure}

@@

\caption{\label{tab:pt2binstable}Correlations between human-coded and model-predicted body position durations using 10-minute bins during the distal comparison period. Correlations are provided within each posture and overall, and computed separately using group and individual models with and without outlier participants.}

@@

\caption{Full-day timelines for each individual recording session, split by (A) younger infants and (B) older infants. Each participant's timeline shows a stacked bar graph with the proportion of time spent in each of the body position for every 5-minute period throughout the day, based on group model predictions of body position. The x-axis shows time of day. Caregiver-reported naps are marked by gray bars; blank gaps indicate caregiver-reported times that the sensor garment was removed for diaper changes, baths, or trips out of the house.}\label{fig:timelines}
\end{figure}

@@

\caption{Age differences in daily body position predicted from (A) group models and (B) individual models. Each circle represents one full-day recording session's proportion of time in each body position (y-axis) against age in months (x-axis). Black lines indicate best fit regression lines indicating decreases in supine time and increases in sitting and upright time with age.}\label{fig:age}
\end{figure}
